How to Analyze Build Logs --- A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-24 tse-paper/content.tex
Finish intro (left some more TODOs)

However, manual analysis of build logs does not
scale to the number of build logs and amount of information that such
sophisticated applications require.
Moreover, being able to
automatically extract relevant pieces of information from build logs
can also help
developers debug a broken build more quickly without having to browse
the entire build log.

@@

Despite its central role for helping developers and for enabling
applications for on-ward processing of build logs,
the process of how to
automatically analyze build logs has thus far received little
attention as a separate research topic, bringing us to the first
research question:
\begin{simplebox}[minipage boxed title*=-5cm]{\textbf{Research Question
1}}
What are the state-of-the-research techniques to analyze build logs?
\end{simplebox}

To answer this question, we start the investigation by performing an
extensive literature study (\Cref{fig:overview}).
It shows that
less than half of the 61 works which rely on build log analysis
describe in detail how they performed it.
Even fewer
made their implementation available.
The literature
survey identifies various strategies to
analyze build logs such as:
custom built parsers, regular expressions, manual inspection,
natural language processing, information retrieval or searching
for keywords.
%Using 1) a set of
% manually curated regular expressions,
% similar to TravisTorrent~\cite{beller2017oops},
% 2) information
% retrieval techniques,
% % todo: Can you shortly explain the technique here?
% and 3) a list of search keywords
% to identify regions of interest in the build log.

Even though these three approaches to analyze
build logs have different strengths and weaknesses, the literature
survey shows that their use often seems arbitrary.
As a consequence,
in the second research question we ask:

@@

To answer this question, in the second step of the study
(\Cref{fig:overview}), we choose three promising
techniques to analyze build logs by retrieving substrings
(chunks) that describe a specific information.

We build and compare three
implementations that each stand prototypical for one of the identified
strategies to analyze build logs:

\begin{itemize}
  \item \textbf{Program Synthesis by Example (referred to as PBE)}
  Based on manually supplied example pairs of build logs and chunks,
  PBE automatically synthesizes
  a regular expression extracting the given chunks within the build logs.
  % these second sentences about what the techniques stand for could
  % also be merged to a paragraph after the itemize
  PBE mimics regular expressions and parsers while
  simplifying their development.
  \item \textbf{Common Text Similarity (CTS)}
  Text similarity is a common information retrieval technique.
  CTS selects lines of a log which are
  most similar to the lines present in manually supplied example chunks.
  As a common information retrieval technique, our text similarity
  approach stands for the machine learning approaches identified in
  the literature survey.
  \item \textbf{Keyword Search (KWS)}
  KWS finds fitting passages by searching the whole
  text of a build log for
  the occurrence of specific trigger words.
  Keyword Search was explicitly mentioned by several articles in the
  literature survey.
\end{itemize}

In the third step (\Cref{fig:overview}),
we conduct an empirical study to gauge the performance
of these chunk retrieval techniques in
real-world projects under a variety of conditions.
Our study is based on the \emph{LogChunks} data
set~\cite{brandt2020logchunks}, which encompasses build logs from
80 popular open-source projects and a broad range of different
build tools and programming languages.

We develop a set of guidelines (step 4, \Cref{fig:overview})
on how to choose a
technique given a goal.
We recommend PBE for use cases where the desired information is always
represented in the same structural way and high confidence in
precision and recall of the chunk retrieval is required.
The results PBE produces are suited for automatic on-ward processing.
CTS is
well-suited when the representation of the desired information varies
slightly and the output of the chunk retrieval is further processed by
a human.
In cases where the textual representation of the desired
information in the log is unpredictable or varies greatly, KWS seems
to be the best choice.
However, its low precision---it extracts a
context of multiple lines around a finding---makes it generally
unsuited for automatic on-ward processing.
Instead, KWS requires a human
to further inspect and interpret the output chunk.
A hybrid-solution combining the three techniques might be the best
all-round solution.
Finally, we give a roadmap to guide future research in the field
of build log analysis.

In short, this article contributes
\begin{itemize}
\item the first systematic study of the emerging field of build log
analysis.
\item three prototypical open-sourced implementations of
promising chunk retrieval techniques.
\item a large empirical study comparing the three techniques.
\item systematic guidelines on when to use which technique.
% TODO: is the future work good enough to put in contributions?
\item a detailed replication package~\cite{brandt2020chunk-replication}.
\end{itemize}

@@

This lead us to the following research questions:
\begin{simplebox}[attach boxed title to top center={yshift=-6mm}]
{\textbf{RQ1:} What is the state-of-the-research on analyzing build logs?}
% TODO (resolved): All of these are passive
% TODO Swap RQ 1.2 and RQ 1.1 (logical order?)
\begin{itemize}[leftmargin=2cm]
  \item[\textbf{RQ1.1:}] Which information in build logs is of interest?
  \item[\textbf{RQ1.2:}] Which extraction techniques exist?
  \item[\textbf{RQ1.3:}] Which kinds of build logs are subject to analysis?
\end{itemize}
\end{simplebox}

@@

\raisebox{0.8mm}{Information Retrieval} &
\raisebox{0.8mm}{
\cite{hassan2017change,lindqvist2019detection,ren2018automated}
} &
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_3.pdf}
\\

\raisebox{0.8mm}{Analysis} &
\raisebox{0.8mm}{
\cite{sulir2016quantitative,haghighatkhah2018test,durieux2019critical}
} &
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_3.pdf}
\\

@@

% TODO QUESTION
% shouldn't it also better be in threats to validity?
% it is as well.
%back for my thesis you said we should put it
% directly here to make it more prominent avoid angry critics
% I think it's good here
% but it's also self-contained so feel free to comment it out
\textbf{Caveat!}
This is a preliminary hypothesis based on the results
from our comparison study.
The recommendations could therefore be
influenced by idiosyncrasies of our specific implementation of the
chunk retrieval techniques as well as the \emph{LogChunks}
data set.

>>>>

However, manual analysis of build logs does not
scale to the number of build logs and amount of information that such
sophisticated applications require.
Moreover, being able to
automatically extract relevant pieces of information from build logs
can help
developers debug a broken build more quickly without having to browse
the entire build log.

@@

Despite its central role in helping developers and in enabling
novel, on-ward processing applications,
the process of how to
automatically analyze build logs has thus far not been systematically overviewed as a research topic, bringing us to the first
research question:
\begin{simplebox}[minipage boxed title*=-5cm]{\textbf{Research Question
1}}
What is the state-of-the-research on analyzing build logs?
\end{simplebox}

To answer this question, we start our investigation---pictured in \Cref{fig:overview}---with an
extensive literature study (step 1).
The literature
survey identifies various strategies to
analyze build logs, among which are developing
custom parsers, using regular expressions, doing manual inspections,
natural language processing, information retrieval, or searching
for keywords.
Unfortunately, our survey also shows that
less than half of the 61 works which rely on build log analysis
describe it in sufficient detail and that
% TODO fill-in X
only X independent implementations are available, one of which stems from the second author of this paper. Out of these implementations, all are based on either a custom parser or regular expressions.
%Using 1) a set of
% manually curated regular expressions,
% similar to TravisTorrent~\cite{beller2017oops},
% 2) information
% retrieval techniques,
% % todo: Can you shortly explain the technique here?
% and 3) a list of search keywords
% to identify regions of interest in the build log.

%% \begin{itemize}
%%   \item \textbf{(referred to as PBE)}
%%   Based on manually supplied example pairs of build logs and chunks,
%%   PBE automatically synthesizes
%%   a regular expression extracting the given chunks within the build logs.
%%   % these second sentences about what the techniques stand for could
%%   % also be merged to a paragraph after the itemize
%%   PBE mimics regular expressions and parsers while
%%   simplifying their development.
%%   \item \textbf{Common Text Similarity (CTS)}
%%   Text similarity is a common information retrieval technique.
%%   CTS selects lines of a log which are
%%   most similar to the lines present in manually supplied example chunks.
%%   As a common information retrieval technique, our text similarity
%%   approach stands for the machine learning approaches identified in
%%   the literature survey.
%%   \item \textbf{Keyword Search (KWS)}
%%   KWS finds fitting passages by searching the whole
%%   text of a build log for
%%   the occurrence of specific trigger words.
%%   Keyword Search was explicitly mentioned by several articles in the
%%   literature survey.
%% \end{itemize}

% TODO caro: maybe convert this into a proper ordered list?
As a consequence, we took it upon ourselves (step 2) to provide implementations of the three most promising automatic build log analysis techniques we observed in the literature survey:
a technique 1) able to automatically synthesize regular expressions, called \emph{Program Synthesis by Example (PBE)}, 2) based on natural language processing, doing \emph{Common Text Similarity (CTS)}  and 3) inspired by information-retrieval employing \emph{Keyword Search (KWS)}.
Even though these three approaches to analyze
build logs have different strengths and weaknesses, the literature
survey shows that their use often seems arbitrary.
As a consequence,
in the second research question we ask:

@@

To answer this question, in the third step  of \Cref{fig:overview},
we conduct an empirical study assessing the performance
of these chunk retrieval techniques in
real-world projects under a variety of conditions.
Our study is based on the \emph{LogChunks} data
set~\cite{brandt2020logchunks}, which encompasses build logs from
80 popular open-source projects and a broad range of different
build tools and programming languages. Our results show that there is no technique that in general outperforms the others. However, 
% TODO 1-2 sentences about results, inclduing some accuracy measures

Finally, due to this ambiguity, we develop a set of guidelines (\Cref{fig:overview}, step 4)
on how to choose the most suitable build log analysis
technique for the task at hand.
We recommend PBE for use cases where the desired information is always
represented in the same structural way and high confidence in
precision and recall of the chunk retrieval is required.
The results PBE produces are suited for automatic on-ward processing.
CTS is
well-suited when the representation of the desired information varies
slightly and the output of the chunk retrieval is further processed by
a human.
In cases where the textual representation of the desired
information in the log is unpredictable or varies greatly, KWS seems
to be the best choice.
However, its low precision---it extracts a
context of multiple lines around a finding---makes it generally
unsuited for automatic on-ward processing.
Instead, KWS requires a human
to further inspect and interpret the output chunk.
A hybrid-solution combining the three techniques might be the best
all-round solution.
Finally, we give a roadmap to guide future research in the field
of build log analysis.

In short, this article contributes
\begin{itemize}
\item the first systematic study of the emerging field of build log
analysis.
\item three prototypical open-sourced implementations of
promising chunk retrieval techniques.
\item a large empirical study comparing the three techniques.
\item systematic guidelines on when to use which technique.
\item a detailed research map to inspire future contributions to the field.
\item a replication package~\cite{brandt2020chunk-replication}.
\end{itemize}

@@

This lead us to the following research questions:
\begin{simplebox}[attach boxed title to top center={yshift=-6mm}]
{\textbf{RQ1:} What is the state-of-the-research on analyzing build logs?}
% TODO (resolved): All of these are passive
% TODO (resolved) Swap RQ 1.2 and RQ 1.1 (logical order?)
\begin{itemize}[leftmargin=1cm]
  \item[\textbf{RQ1.1:}] What information is extracted from build logs?
  \item[\textbf{RQ1.2:}] Which extraction techniques exist?
  \item[\textbf{RQ1.3:}] Which kinds of build logs are subject to
  analysis?
\end{itemize}
\end{simplebox}

@@

\raisebox{0.8mm}{Information Retrieval} &
\raisebox{0.8mm}{
\cite{hassan2017change,lindqvist2019detection,ren2018automated}
} &
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_3.pdf}
\\
% TODO: We need to mention in the text describing the Table what we mean by ``Analysis''
\raisebox{0.8mm}{Analysis} &
\raisebox{0.8mm}{
\cite{sulir2016quantitative,haghighatkhah2018test,durieux2019critical}
} &
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_3.pdf}
\\

@@

% TODO QUESTION
% shouldn't it also better be in threats to validity?
% it is as well.
%back for my thesis you said we should put it
% directly here to make it more prominent avoid angry critics
% I think it's good here
% but it's also self-contained so feel free to comment it out

% TODO we can't just have a paragraph in a scientific paper start with Caveat, I think ;-) Also, we should say why we THINK this is not the case. Otherwise we weaken our work unduly
\textbf{Caveat!}
This is a preliminary hypothesis based on the results
from our comparison study.
The recommendations could therefore be
influenced by idiosyncrasies of our specific implementation of the
chunk retrieval techniques as well as the \emph{LogChunks}
data set.

### 2020-04-24 tse-paper/paper.tex
Finish intro (left some more TODOs)

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
Build logs are the textual by-products that automatic software build processes
create, often as part of their Continuous Integration (CI)
pipeline.
Build logs can serve as a paramount source of
information for developers to understand and debug depdency, compilation, or
test failures.
Manually
extracting the important chunks of information in a build log, though, is akin to
finding a needle in a haystack.
Recently, practitioners and researchers have started attempts to partly
automate
this time-consuming activity, to not only ease
developers' task of finding this needle but also to enable a novel class of
automated on-ward processing applications.
In this paper, we give a systematic overview of the emerging field of build log analysis.
In a large-scale literature mapping study, we first survey and categorize the existing methods to
extract information chunks from build logs.
In the second part of the paper, we develop prototypical implementations of the three most promising
techniques we found, namely program synthesis
by example, textual similarity, and keyword search.
Finally, we evaluate the techniques in an empirical study on
the \emph{LogChunks} data set, which comprises 797 build logs 
from a wide range of different projects and programming languages.
Our findings show that none of the three techniques in general outperforms
the others, but rather, that each technique has its respective strengths
and weaknesses.
We discuss under which circumstances each technique performs best
and provide a recommendation scheme for when developers or researchers might 
use which technique.
We close with a roadmap of important open problems in the
emerging field of
build log analysis.
\end{abstract}

>>>>

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
Build logs are the textual by-products that automatic software build processes
create, often as part of their Continuous Integration (CI)
pipeline.
Build logs can serve as a paramount source of
information for developers to understand and debug depdency, compilation, or
test failures.
Manually
extracting the important chunks of information in a build log, though, is akin to
finding a needle in a haystack.
Recently, practitioners and researchers have started attempts to partly
automate
this time-consuming activity, to not only ease
developers' task of finding this needle but also to enable a novel class of
automated on-ward processing applications.
In this paper, we give a systematic overview of the emerging field of build log analysis.
In a large-scale literature mapping study, we first survey and categorize the existing methods to
extract information chunks from build logs.
In the second part of the paper, we develop prototypical implementations of the three most promising
techniques we found, namely program synthesis
by example, textual similarity, and keyword search.
Finally, we evaluate the techniques in an empirical study on
almost 800 build logs 
from a wide range of different projects and programming languages.
Our findings show that none of the three techniques in general outperforms
the others, but rather, that each technique has its respective strengths
and weaknesses.
We discuss under which circumstances each technique performs best
and provide a recommendation scheme for when developers or researchers might 
use which technique.
We close with a roadmap of important open problems in the
emerging field of
build log analysis.
\end{abstract}
